Standalone Tactics using OpenTheory

### 2012-02-08 paper.tex
work on sec 2.2

\newcommand{\OpenTheory}{OpenTheory\xspace}
\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\ie}{i.e.\xspace}

@@

\subsection{Isolating tactics from their home systems}
There are two approaches to obtaining a standalone tactic: either write one directly, or extract an existing tactic from a proof assistant. 
We have experimented with the second approach, taking tactics from HOL4 and from HOL Light.
The procedure is reasonably lightweight, but less flexible than writing a standalone tactic directly.
The key requirements of the proof assistant are:
\begin{enumerate}
\item
\label{req:io} an \OpenTheory article file reader and writer,
\item
\label{req:log} a facility for recording proofs and converting them to the \OpenTheory logical kernel, and,
\item
\label{req:ckpt} a facility for checkpointing the proof assistant, or for extracting a standalone executable encompassing the tactic functionality.
\end{enumerate}
<first two items are the most work, but have the additional benefit of allowing import/export of theories in opentheory repositories, as mentioned above. so: to use standalone tactics you also get to use opentheory packages, but if you can make yourself into a standalone tactic you can also export opentheory packages.>
<in HOL4 and HOL Light, item 2 is accomplished by augmenting the internal theorem type with a type of proofs, to track primitive inferences used. this step is made easy by LCF architecture: could be very difficult in systems that don't already generate proofs somehow, or with significantly different kernel (cf isabelle)>
<either use Polyml.export or checkpointing to capture a function that reads a goal article, runs the tactic (recording the proof), and writes the proof article back>

>>>>

\newcommand{\OpenTheory}{OpenTheory\xspace}
\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\ie}{i.e.\xspace}
\newcommand{\eg}{e.g.\xspace}

@@

\subsection{Isolating tactics from their home systems}
There are two approaches to obtaining a standalone tactic: either write one directly, or extract an existing tactic from a proof assistant. 
We have experimented with the second approach, taking tactics from HOL4 and from HOL Light.
The procedure is reasonably lightweight, but less flexible than writing a standalone tactic directly.
The key requirements of the proof assistant are:
\begin{enumerate}
\item
\label{req:io} an \OpenTheory article file reader and writer,
\item
\label{req:log} a facility for recording proofs and converting them to the \OpenTheory logical kernel, and,
\item
\label{req:ckpt} a facility for checkpointing the proof assistant, or for extracting a standalone executable encompassing the tactic functionality.
\end{enumerate}

Just as the requirements of a system to use standalone tactics also enable it to import \OpenTheory packages in general, the requirements to create standalone tactics (specifically \ref{req:io} and \ref{req:log}) also enable a system to create and export \OpenTheory packages.

Requirement~\ref{req:log} above poses the highest barrier: a standalone tactic must record proofs at a level of detail sufficient to create proofs of the same theorems using the \OpenTheory kernel.
It helps if the system containing the tactic uses the LCF architecture, that is, theorems can only be created by a small number of primitive inference rules.
In HOL4 and HOL Light, we met Requirement~\ref{req:log} by augmenting the internal theorem type with a type of proofs, to track primitive inferences used.
Then we express each primitive inference as a derived rule above the \OpenTheory kernel.
In many cases, there is a direct counterpart (\eg reflexivity of equality is primitive in all three kernels).
But, for example, HOL4's primitive rule for definition by specification must be emulated by a definition using Hilbert choice in \OpenTheory.
<could be difficult in system with significantly different kernel, eg isabelle>

<either use Polyml.export or checkpointing to capture a function that reads a goal article, runs the tactic (recording the proof), and writes the proof article back>
